---
title: "Wikimedia Clickthrough Analysis"
author: "Jim Neville"
date: "May 26, 2018"
output: 
  html_document: 
    keep_md: yes
---

## Executive Summary

The event logs contained in `events_log.csv` include data regarding Wikimedia Searches from 3/1/16 through 3/8/16. The data suggest that the treatment in Group A tends to coincide with a greater average clickthrough rate, as well as a longer average session length. Additionally, it appears that the differences between the  Group A and Group B treatments had no significant effect on the zero results rate.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load required packages and read in dataset
library(tidyverse)
library(knitr)

df <- data.table::fread('events_log.csv')
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# convert variable datatypes
df <- df %>% mutate_at(.vars = c('session_id', 'group', 'action'), .funs = as.factor)
df$timestamp <- lubridate::ymd_hms(df$timestamp)
df$date <- lubridate::as_date(df$timestamp)
df$date <- as.factor(df$date)
```


### Daily Overall Clickthrough Rate

The overall clickthrough rate remained steady over this period. The minimum clickthrough rate (24.31%) was recorded on 3/2/18, while the maximum clickthrough rate (25.75%) was recorded on 3/4/18. 

```{r echo=FALSE}
# Create clickthrough rate dataframe
searches <- df %>% 
  filter(action == 'searchResultPage') %>% 
  group_by(date) %>% 
  summarise(searches = n())

clickthroughs <- df %>% 
  filter(action == 'visitPage' & !is.na(result_position)) %>% 
  group_by(date) %>% 
  summarise(clickthroughs = n())

clickthrough_rate <- searches %>% 
  left_join(clickthroughs, by = 'date') %>% 
  mutate(clickthrough_rate = clickthroughs/searches)
```


```{r echo=FALSE}
# Plot Daily clickthrough rate
vis_theme <- ggthemes::theme_tufte(ticks=F)

clickthrough_rate %>% ggplot() +
  geom_line(aes(as.factor(date), clickthrough_rate, group=1, color = "red"), size = 1.5) +
  geom_point(aes(as.factor(date), clickthrough_rate, group=1, color = "red"), size = 2.2) +
  vis_theme +
  labs(title = 'Daily Overall Clickthrough Rate', x = 'Date', y = 'Clickthrough Rate') +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0,0.3)) 
```


### Daily Clickthrough Rate - AB Test

Over this period, it is clear that the Group A treatment led to a much greater clickthrough rate compared with the Group B treatment. The Group A treatment had a maximum clickthrough rate of 30.62% and a minimum clickthrough rate of 28.43%, while the Group B treatment had a maximum clickthrough rate of 17.64% and a minimum clickthrough rate of 12.05%. Moreover, a paired t-test revealed a statistically significant difference between the two treatment groups (t = 15.41, p < 0.01).

```{r echo=FALSE}
# Create clickthrough rate AB test dataframe
searches_ab <- df %>%
  filter(action == 'searchResultPage') %>% 
  group_by(date, group) %>% 
  summarise(searches = n())

clickthroughs_ab <- df %>% 
  filter(action == 'visitPage'& !is.na(result_position)) %>% 
  group_by(date, group) %>% 
  summarise(clickthroughs = n())

clickthrough_rate_ab <- clickthroughs_ab %>% 
  left_join(searches_ab, by = c('date', 'group')) %>% 
  mutate(clickthrough_rate = clickthroughs/searches)
```


```{r echo=FALSE}
# Plot daily clickthrough rate - AB test
tmp <- clickthrough_rate_ab
names(tmp)[2] <- "Group"
tmp <- tmp %>% mutate(Group = case_when(Group == "a" ~ "A", 
                                        Group == "b" ~ "B"))

tmp %>%
  ggplot() +
  geom_line(aes(as.factor(date), clickthrough_rate, group = Group, color = Group), size = 1.5) +
  geom_point(aes(as.factor(date), clickthrough_rate, group = Group, color = Group), size = 2.2) +
  vis_theme +
  labs(title = 'Daily Clickthrough Rate - AB Test', x = 'Date', y = 'Clickthrough Rate') +
  theme(legend.position = 'top', plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0,0.4))
```


```{r echo=FALSE}
# Create dataframe for t test and run t test
ctr_a_tmp <- clickthrough_rate_ab %>% filter(group == "a") %>% mutate(ctr_grp_a = clickthrough_rate)
ctr_b_tmp <- clickthrough_rate_ab %>% filter(group == "b") %>% mutate(ctr_grp_b = clickthrough_rate)
ctr_t_test <- ctr_a_tmp %>% left_join(ctr_b_tmp, by = "date")

ctr_ab_t_test <- t.test(ctr_t_test$ctr_grp_a, ctr_t_test$ctr_grp_b, alternative = "greater", paired = TRUE, conf.level = 0.99)

broom::tidy(ctr_ab_t_test)
```


### Results Users Try First

Overall, the majority of all searches (53%) result in the user visiting the webpage listed first in the search results, with the liklihood that the user will click on the result decreasing as the result positiion moves farther away from 1. 

```{r echo=FALSE}
# Plot result positions clicked most frequently
visit_vis <- df %>% filter(action == 'visitPage' & !is.na(result_position) & result_position < 11) %>% 
  ggplot(aes(as.factor(result_position), fill = 'red')) + 
  geom_bar()

visit_vis +
  vis_theme +
  labs(title = 'Result Positions Clicked Most Frequently', x = 'Result Position', y = 'Frequency') +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::comma,limits = c(0, 25000))
```

This trend remained steady over the 8-day period from 3/1/16 to 3/8/16.

```{r echo=FALSE}
# Plot result positions clicked most frequently by date
visit_vis +
  facet_wrap(~ date) +
  vis_theme +
  labs(title = 'Result Positions Clicked Most Frequetly by Date', x = 'Result Position', y = 'Frequency') +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::comma,limits = c(0, 3500))
```

### Daily Overall Zero Results Rate

The Daily Overall Zero Results Rate remained faily steady over this time period, with the maximum Zero Results Rate (24.03%) occuring on 3/8/18 and the minimum Zero Results Rate (21.38%) occuring on 3/6/18.

```{r echo=FALSE}
# Create zero results rate dataframe
search_nr <- df %>% 
  filter(action == 'searchResultPage' & n_results == 0) %>% 
  group_by(date) %>% summarise(no_results = n())

search_r <- df %>% 
  filter(action == 'searchResultPage' & n_results > 0) %>% 
  group_by(date) %>% summarise(results = n())

daily_zero_results_rate <- search_nr %>% 
  left_join(search_r, by = "date") %>% 
  mutate(zero_results_rate = no_results/results)
```


```{r echo=FALSE}
# Plot daily zero results rate
daily_zero_results_rate %>% 
  ggplot(aes(as.factor(date), zero_results_rate, color = "red")) +
  geom_line(group = 1, size = 1.5) +
  geom_point(group = 1, size = 2.2) +
  vis_theme +
  labs(title = 'Daily Overall Zero Results Rate', x = 'Date', y = 'Zero Results Rate') +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0, 0.25))
```

### Daily Zero Results Rate - AB Test

The Daily Zero Results Rate does not appear to vary a great deal between the group treatments. Moreover a one-tailed paired t-test revealed that there is not a statistically significant difference between the Zero Results Rate for the two treatment groups (t = -0.86, p = 0.21) 

```{r echo=FALSE}
# Create daily zero results rate AB test dataframe
search_nr_ab <- df %>% 
  filter(action == 'searchResultPage' & n_results == 0) %>% 
  group_by(date, group) %>% 
  summarise(no_results = n())

search_r_ab <- df %>% 
  filter(action == 'searchResultPage' & n_results > 0) %>% 
  group_by(date, group) %>% 
  summarise(results = n())

daily_zero_results_rate_ab <- search_nr_ab %>% 
  left_join(search_r_ab, by = c("date", "group")) %>% 
  mutate(zero_results_rate = no_results/results)
```


```{r echo=FALSE}
# Plot daily zero results rate ab test
tmp <- daily_zero_results_rate_ab
names(tmp)[2] <- "Group"
tmp <- tmp %>% mutate(Group = case_when(Group == "a" ~ "A",
                                        Group == "b" ~ "B"))

tmp %>% 
  ggplot(aes(as.factor(date), zero_results_rate, group = Group, color = Group)) +
  geom_line(size = 1.5) +
  geom_point(size = 2.2) +
  vis_theme +
  labs(title = 'Daily Zero Results Rate - AB Test', x = 'Date', y = 'Zero Results Rate') +
  theme(legend.position = 'top', plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0,0.3))
```


```{r echo=FALSE}
# Create dataframe for t test and run t test
zrr_a_tmp <- daily_zero_results_rate_ab %>% filter(group == "a") %>% mutate(zrr_grp_a = zero_results_rate)
zrr_b_tmp <- daily_zero_results_rate_ab %>% filter(group == "b") %>% mutate(zrr_grp_b = zero_results_rate)
zrr_t_test <- zrr_a_tmp %>% left_join(zrr_b_tmp, by = "date")

zrr_ab_t_test <- t.test(zrr_t_test$zrr_grp_a, zrr_t_test$zrr_grp_b, alternative = "less", paired = TRUE, conf.level = 0.95)

broom::tidy(zrr_ab_t_test)
```

### Session Length

The searches in the Group A treatment had a mean session length of 102 seconds, while the Group B treatment's mean was much less at 29.1 seconds.

```{r echo=FALSE}
## Create session length dataframe
session_length <- df %>% 
  replace_na(replace = list(checkin = 0)) %>% 
  group_by(session_id) %>%
  summarise(session_length = max(checkin) - min(checkin))

df$group <- as.character(df$group)

group_by_session <- df %>% 
  group_by(session_id) %>% 
  summarise(group = max(group))

session_length_group_cor <- group_by_session %>% 
  left_join(session_length, by = "session_id")

session_length_group_cor <- session_length_group_cor %>% 
  group_by(group) %>% summarise(mean_sl = mean(session_length))
```


```{r echo=FALSE}
# Plot average session length by group
ggplot(session_length_group_cor, aes(group, mean_sl, group = group, color = "red")) +
  geom_line(group = 1, size = 1.5) +
  geom_point(group = 1, size = 2.2) +
  vis_theme +
  labs(title = 'Average Session Length by Group', x = 'Group', y = 'Average Session Length (in seconds)') +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0,120)) +
  scale_x_discrete(labels = c("A", "B"))
```

